---
title: "Take-home assignment for: Data Analyst/Programmer - Thera Business Inc"
author: "Jiahao Ye"
date: "2023-02-02"
output:
  html_document:
    df_print: paged
---

### Task#1: Load the dataset in R and label the dataset as "test_thera"

```{r}
# Load readxl package
library("readxl")
# Load the stringr package
library(stringr)

# load the data and rename
test_thera = read_excel("D:/dataset/Research-Analyst_16SEPT2021.xlsx")

```

### Task#2: Present the names of variables

```{r}
# present a list of variable names in the form
names(test_thera)

```

The variables are shown in the above. There are in total 18 columns/variables

### Task#3: Find the number of rows and columns in this dataframe

```{r}
number_of_rows = nrow(test_thera)
number_of_cols = ncol(test_thera)

str_glue("This dataframe has {number_of_rows} number of rows and {number_of_cols} number of columns.")

```

### Task#4: Show the last 6 rows of "age":

```{r}
tail(test_thera[["age"]],6)

```

The last 6 rows of column 'age' are: 21,43,39,37,45,62

### Task#5: Replace column name from "gender" to "sex"; replace last column to "blood pressure"

```{r}
# replace gender
colnames(test_thera)[which(names(test_thera) == "gender")] = "sex"
# replace 'blood pressure'
colnames(test_thera)[which(names(test_thera) == "systolic bp preoperative")] = "blood pressure"
```

### Task#6: Replace the values in 'income bracket': 1-\> \<10,000....

```{r}

# first, modify the variable type in column 'income bracket'
test_thera$`income bracket` = as.character(test_thera$`income bracket`)

# loop through the age column from top to bottom and replace value
for (i in 1:nrow(test_thera)) {
  
  if (test_thera[i,"income bracket"] == 1) {test_thera[i,"income bracket"] =  "<10,000"}
  else if (test_thera[i,"income bracket" ] == 2) {test_thera[i,"income bracket"] = "10,000 to 20,000"}
  else if (test_thera[i,"income bracket" ] == 3) {test_thera[i,"income bracket"] = "20,001 to 30,000"}
  else if (test_thera[i,"income bracket" ] == 4) {test_thera[i,"income bracket"] = "30,001 to 40,000"}
  else if (test_thera[i,"income bracket" ] == 5) {test_thera[i,"income bracket"] = ">40,001"}
}

```

Now the values in 'income bracket' has been replaced

### Task#7: Run the first 6 rows of "income bracket"

```{r}
head(test_thera$`income bracket`,6)

```

The first 6 values from income column are shown as above

### Task#8: Do a descriptive analysis on: "duration of operation" and "systolic bp preoperative". Run an appropriate MICE procedure to impute for missing values

first check if there exists any missing values in these 2 columns

```{r}
sum(is.na(test_thera$`duration of operation`))
sum(is.na(test_thera$`blood pressure`))
```

there are 12 missing values for 'duration of opeartion' and 34 missing values for 'systolic bp preoperative'

use MICE procedure to impute for missing value: MICE stands for Multivariate Imputation by Chained Equation algorithm and it is an algorithm used to fill in the blanks. It simply uses values in other columns to predict the missing value

```{r}
# first import the mice package
library(VIM)

# analyze the missing values:
aggr_plot <- aggr(test_thera, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(test_thera), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
```

As can be seen from the visualization above, only 'duration of opeartion' and 'systolic bp preoperative' has missing values. Missing values are of about 3.3% and 1.1% respectively.

After knowing the nature of the missing data, then the built-in mice() function should be used to compute the missing value

```{r}

# there is an error when I try to run mice() on the test_thera dataframe; so I just create a deep copy(with different memory location) of the test_thera dataframe and run mice() on it.
test_thera_copy = data.frame(test_thera);

# turn the non-continuous variable into categorical variables:
names <- c("type.of.surgery","use.of.postoperative.drain", "entry.of.paranasal.sinus","CSF.leak","diabetes.mellitus" ,"SSI","discharge.status","glucorticoids","lumbar.drainage","income.bracket","hospital.code","sex")
test_thera_copy[,names] <- lapply(test_thera_copy[,names] , factor)
str(test_thera_copy)
```

It can be confirmed that variables such as gender, type of surgery etc has been converted into categorical variables

We can start the MICE imputation process now

```{r}
library(mice);
init = mice(test_thera_copy, maxit=0) 
meth = init$method
predM = init$predictorMatrix

# remove the NA variable and not include that as a predictor
predM[, c("blood.pressure","duration.of.operation")]=0

# specify the method for imputation
meth[ c("blood.pressure","duration.of.operation")]="norm"
```

start the imputation process

```{r}
set.seed(103)
imputed = mice(test_thera_copy, method=meth, predictorMatrix=predM, m=5)
```

```{r}
imputed <- complete(imputed)

sapply(imputed, function(x) sum(is.na(x)))

```

It can be confirmed that all the missing values has been filled. The dataframe that I will be using for the next questions will be 'imputed'

Perform a descriptive analysis for: 'duration of operation' and 'systolic bp preoperative'

```{r}
# use the summary function to find some descriptive data
print("-------duration of operation-----")
summary(imputed$duration.of.operation)
print("-------blood pressure-----")
summary(imputed$blood.pressure)
```

Draw histograms and boxplots: find out how the data is distributed

-   Observation: not very close to a normal distribution, more like a uniform distribution

```{r}
library('ggplot2')
# histogram:
hist(imputed$duration.of.operation)
hist(imputed$blood.pressure)
# boxplot:
boxplot(imputed$duration.of.operation)
boxplot(imputed$blood.pressure)
# QQPlot
qqnorm(imputed$duration.of.operation)
qqnorm(imputed$blood.pressure)
```

### Task#9: Perform descriptive analysis on 5 other variables of your choice, and develop some graph

I am interested in seeing the following:

-   age distribution
-   income group distribution
-   the discharge.status distribution across different income group
-   numberofdays in ICU distirbution across different hispital
-   how does blood pressure differs if lumbar.draingage exists or not

First find out the age and income group distribution

-   observation: age range is mostly 20-70 years old; income distribution is quite even.

```{r}
hist(imputed$age, main="Age distribution")
pie(table(imputed$income.bracket), col=grey.colors(3), main="Income distribution")
```

Find out the discharge.status distribution across different income group

-   observation: there is no significant difference of alive-dead status across different income group

```{r}
# I will simply use a mosaicplot to show 
mosaicplot(income.bracket~discharge.status,data=imputed,col=c("Blue","Red"))

```

Find out numberofdays in ICU distirbution across different hispital

-   observation: At hospital#5, the count of 2-day.in.ICU is a little higher compared to other hospitals. The occupations of ICUs in different hospitals are about the same.

```{r}
# mosaicplot(hospital.code~days.in.the.NICU,data=imputed,col=c("Blue","Red"))
ggplot(data = imputed)+geom_bar(aes(x=hospital.code, fill=as.factor(days.in.the.NICU))) + ggtitle(label="The distribution of days in ICU across different hospital")+ theme_bw()
```

Find lumbar.drainage vs blood pressure

-   Looking from the graph, the use of lumbar drainage does not have a major impact on blood pressure

```{r}
# draw 2 boxplots side by side
library(ggplot2)
ggplot(imputed, aes(lumbar.drainage, y=blood.pressure)) + geom_boxplot() + theme_bw()
```

Task#10: Low-risk versus High-risk

-   1\~2 days: low-risk, more than 2 days: high-risk
-   compare the characteristics btw patients who are "low-risk" compared to "high-risk"

First prepare the data by creating a new column 'risk.level', with categoraical variable:0 and 1

```{r}
# create a new column called risk level; 0->'low-risk' 1->'high-risk'
imputed$'risk.level' = as.factor(ifelse(imputed$days.in.the.NICU <= 2, 0, 1))

# find some descriptive statistics about the risk-lvel column
print("---count---")
summary(imputed$risk.level)

# use the aggregate function to find the mean of continuous variable for each categorical variable
a <- aggregate(age ~ risk.level, data=imputed, FUN=mean)
barplot(a$age, names.arg=a$risk.level)

b <- aggregate(duration.of.operation ~ risk.level, data=imputed, FUN=mean)
barplot(b$duration.of.operation, names.arg=b$risk.level)
```

As can be seen from the 2 aggregate plot from above, the average age for high-risky group tends to be slightly higher; the average duration.of.operation for high-risky group tends to be lower

### Task#11: Esitimate the likelihood for being low-risk or high-risk

-   Perform some univariate regression first: run univariate regression on: days.in.ICU \~ many other variable, and check the R value for statistical significance

-   Then run a multivariate regression; add the categorical variables into the regression model as well

```{r}
lm1 <- lm(days.in.the.NICU ~ age, data = imputed)
lm2 <- lm(days.in.the.NICU ~ duration.of.operation, data = imputed)
lm3 <- lm(days.in.the.NICU ~ blood.pressure, data = imputed)

# print the linear regression model result
summary(lm1)
summary(lm2)
summary(lm3)
```

As can be seen from the regression analysis above, the variable:duration.of.operation is statistically insignificant (p value \> 0.1, not significant even at 10% level); the other 2 varibales: age & blood.pressure are significant, and I will incorporate these 2 variables into the multivariate regression model

Run a multivariate regression

```{r}
multivariate.lm = lm(formula = days.in.the.NICU~age+blood.pressure+sex+glucorticoids+diabetes.mellitus+CSF.leak+SSI+use.of.postoperative.drain+lumbar.drainage, data=imputed)

summary(multivariate.lm)
```

As can be seen from the summary above, in the regression model, the following variables are statistically significant

-   age - \<0.001
-   blood.pressure - 0.01
-   male - 0.01
-   SSI - 0.01

Incorporate some other categorical variables into the multivariate regression model:

```{r}
multivariate.lm2 = lm(formula = days.in.the.NICU~age+blood.pressure+sex+SSI+hospital.code+type.of.surgery+income.bracket, data=imputed)

summary(multivariate.lm2)
```

As can be seen from the summary above, there are a couple observations:

-   the hospital that the patient is in have no impact on his/her number of days in ICU
-   the type of surgery that the patient has have no impact on his/her number of days in ICU
-   A higher income level (\>40,001) has negative correlation to his/her number of days in ICU

Run a logistic regression(0 or 1) to find the Likelihood of being low-risk and high-risk

```{r}
## 'risk level' is a categorical variable with 0 and 1 
multi_logit=glm(`risk.level`~age+sex+`diabetes.mellitus`+`CSF.leak`+`duration.of.operation`+`income.bracket`+`blood.pressure`+`type.of.surgery`,data=imputed,family=binomial(link="logit"))
summary(multi_logit)

```

Observations:

-   an older age && being a male && having higher blood pressure increase the chance of being 'high-risk'
-   Other variables does not provide any predictive power because they are statistically insignificant

## END OF DOCUMENT
